Do we need detailed demographic data to forecast plant population responses to climate change?

### 2016-05-25 manuscript/manuscript.Rmd
updates ms with MBH edits

\textsf{\textit{\footnotesize{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322, USA; \textsuperscript{2}U.S. Geological Survey, Colorado Cooperative Fish and Wildlife Research Unit, Department of Fish, Wildlife and Conservation Biology, Colorado State University, Fort Collins, CO 80523-1403, USA; \textsuperscript{3}Department of Fish, Wildlife, and Conservation Biology, Colorado State University, Fort Collins, CO 80523 USA}; \textsuperscript{4}Department of Statistics, Colorado State University, Fort Collins, CO 80523 USA}}

@@

Aggregate measures of population status, rather than individual performance, offer an intriguing alternative for modeling populations [@Clark2004; @Freckleton2011]. Population-level data cannot provide inference about demographic mechanisms, but might be sufficient for modeling future population states, especially since such data are feasible to collect across broad spatial extents [e.g., @Queenborough2011]. The choice between individual and population-level data involves a difficult trade-off: while individual-level data are necessary for mechanistic models, population-level data enable models that can be applied over greater spatial and temporal extents. An open question is how much forecasting skill is lost when we build models based on population rather than individual-level data.

@@

Here, we compare the forecasting skill of statistical and population models based on aggregated, population-level data with the skill of models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We simulated the models to answer two questions motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b].
First, can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
Second, can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data?

Materials and Methods
---------------------
### Study site and data
Our demographic data come from a northern mixed grass prairie at the Fort Keogh Livestock and Range Research Laboratory near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *P. secunda* are cool-season perennial grasses. All species typically begin growth in the early spring, reach maximum growth and flower in early to mid summer (May-June), and disperse seed in mid to late summer (July-September). [LAST SENTENCES GLOSSES OVER LIKELY BIG DIFFERENCES IN PHENOLOGY. JUST DESCRIBE TIMING OF GROWING SEASON INSTEAD, TO SET UP CLIMATE COVARIATES?]

@@

From 1932 to 1945 individual plants were identified and mapped annually in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were distributed among six pastures, each assigned a grazing treatment of light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76 ha/aum) stocking rates (two pastures per treatment). In this analysis we account for potential differences among the grazing treatments, but do not focus on grazing$\times$climate interactions. The annual maps of the quadrats were digitized and the fates of individual plants tracked and extracted using a computer program [@Lauenroth2008; @Chu2014a]. The permanent quadrats have not been relocated, but their distribution in six different pastures means the data represent a broad spatial distribution for the study area. Daily climate data are available for the duration of the data collection period (1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the study site.

We modeled each grass population based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat-level, we data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because measurement error at the individual-level is small [@Chu2014]. Based on these two datasets of 13 year-to-year transitions, we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level we explicitly model three vital rates: growth, survival, and recruitment. At the quadrat level we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1*, ignoring within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Sample sizes for each species and vital rate model are shown in Table 1.

All R code and data necessary to reproduce our analysis is archived on GitHub as release v1.0[^revnote] (http://github.com/atredennick/MicroMesoForecast/releases). That stable release will remain static as a record of this analysis, but subsequent versions may appear if we update this work. We have also deposited the v1.0 release on Dryad (*link here after acceptance*).

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately since they required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data (Chu et al. *in press*) and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

We fit all models using a hierarchical Bayesian approach. The models are fully descibed in Appendix A, so here we focus on the main process and the model likelihood. For the likelihood models, $\textbf{y}^X$ is always the relevant vector of observations for vital rate *X* (*X* = *S*, *G*, *R*, or *P*) for survival, growth, recruitment, or population growth). For example, $\textbf{y}^S$ is a vector of 0's and 1's indicating whether a genet survives from *t* to *t+1*, or not. All model parameters are species-specific, but we omit subscripts for species in model descriptions below to reduce visual clutter. For brevity, we only describe models with climate covariates included, but models without climate covariates are simply the models described below with the climate effects removed.

#### Vital rate models at the individual level
We used logistic regression to model survival probability ($s$) of a genet:

\vspace{-3em}
\begin{align}
\textbf{y}^{S} &\sim \text{Bernoulli}(\textbf{s}) \\
\text{logit}[s(x,\textbf{z}_t,w)] &= \beta_{0,t} + \beta_{s,t}x + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w + \beta_{d,2} (xw)
\end{align}
\vspace{-3em}

where $x$ is the log of genet basal area, $\beta_{0,t}$ is a year specific intercept intercept, $\beta_Q$ is the random effect of quadrat group location, $\beta_{s,t}$ is the year-specific slope parameter for size, **z** is a vector of *i* climate covariates specific to year *t*, $\boldsymbol{\beta}_c$ is a vector of fixed climate effects of length *i*, $\beta_{d,1}$ is the effect of intraspecific crowding experienced by the focal genet ($w$), and $\beta_{d,2}$ is a size by crowding ($xw$) interaction effect.

@@

\begin{equation}
w_{i,t} = \sum_k e^{-\deltad_{ik,t}^{2}}u_{k,t}.
\end{equation}

In the above, $w_{i,t}$ is the crowding that genet _i_ in year _t_ experiences from conspecific neighbors. The spatial scale over which conspecific neighbors exert influence on any genet is determined by $\delta$. The function is applied for all _k_ conspecific genets that neighbor the focal genet at time _t_, and $d_{ik,t}$ is the distance between genet _i_ and conspecific genet _k_. We use regression-specific (survival and growth) $\delta$ values estimated by @Chu2015.

We modeled growth as a Gaussian process describing genet size at time $t+1$ as a function of size at $t$ and climate covariates:

\vspace{-3em}
\begin{align}
\textbf{y}^G &\sim \text{Normal}(\boldsymbol{\mu}, \sigma^2_{x,t}) \\
\mu(x,\textbf{z}_t,w) &= \beta_{0,t} + \beta_{s,t}x + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w + \beta_{d,2} (xw)
\end{align}
\vspace{-3em}

where $\mu(x,\textbf{z}_t,w)$ is log of predicted genet size at time *t* + 1, and all other parameters are as described for the survival regression. We capture non-constant error variance in growth by modeling the variance around the growth regression ($\sigma^2_{x,t}$) as a nonlinear function of predicted genet size:

\vspace{-3em}
\begin{align}
\sigma^2_{x,t} = a \, \text{exp}[b \times \mu(x,\textbf{z}_t,w)]
\end{align}
\vspace{-3em}

where $\mu(x,\textbf{z}_t,w)$ is log of predicted genet size predicted from the growth regression (Eq. 4), and *a* and *b* are constants.

Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. 
Therefore, we model recruitment at the quadrat level.
We assume the number of individuals, $y^{R}$, recruiting at time $t+1$ in quadrat q follows a negative binomial distribution:

@@

\vspace{-3em}
\begin{align}
\lambda_{q,t+1} = C'_{q,t} \, \text{exp}\left(\beta_{0,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d}\sqrt{C'_{q,t}}\right)
\end{align}
\vspace{-3em}

where $C'$ is effective cover ($\text{cm}^2$) of the focal species in quadrat $q$, and all other terms are as in the survival and growth regressions.
Effective cover is a mixture of observed cover ($C$) in the focal quadrat ($q$) and the mean cover across the entire group ($\bar{C}$) of $Q$ quadrats in which $q$ is located:

\vspace{-3em}
\begin{align}
C'_{q,t} = pC_{q,t} + (1-p)\bar{C}_{Q,t}
\end{align}
\vspace{-3em}

where $p$ is a mixing fraction between 0 and 1 that is estimated within the model.

#### Population model at the quadrat level
The statistical approach used to model aggregated data depends on the type of data collected. We have percent cover data, which can easily be transformed to proportion data. An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is [0,1], not including true zeros or ones. However, when we used fitted model parameters from a beta regression in a quadrat-based population model, the simulated population tended toward 100% cover for all species. We therefore chose a modeling approach based on a truncated log-normal likelihood. The model for quadrat cover change from time $t$ to $t+1$ is

\vspace{-3em}
\begin{align}
\textbf{y}^{P} &\sim \text{LogNormal}(\mu(x,\textbf{z}_t), \sigma^2) \text{T}[0,1] \\
\mu(x,\textbf{z}_t) &= \beta_{0,t} + \beta_{s,t}x + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c
\end{align}
\vspace{-3em}

where $\mu(x,\textbf{z}_t)$ is the log of proportional cover in quadrat $q$ at time $t+1$, and all other parameters are as in the individual-level growth model (Eq. 4) except that *x* now represents log of proportional cover. The log normal likelihood includes a truncation (T[0,1]) to ensure that predicted values do not exceed 100% cover.

### Model fitting and statistical regularization
####  Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. Quadrat random effects were also fit hierarchically, with quadrat offsets being drawn from distributions with mean zero and a shared variance term (independent Gaussian priors, Appendix A). Climate effects were modeled as independent covariates whose prior distributions were determined by statistical regularization (see **Statistical regularization: Bayesian ridge regression** below).

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. Such short MCMC chains may surprise readers more familiar with other MCMC samplers (i.e. JAGS or WinBUGS), but the Stan sampler is exceptionally efficient, which reduces the number of iterations needed to achieve convergence. We assessed convergence visually and made sure scale reduction factors for all parameters were less than 1.1. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains to be used as randomly drawn values during population simulation. We report the posterior mean, standard deviation, and 95% Bayesian Credible Intervals for every parameter of each model for each species in Appendix B.

#### Statistical regularization: Bayesian ridge regression
For models with climate covariates, our objective is to model the response of our focal grass species to interannual variation in climate, even if those responses are weak.
Therefore, we avoid selecting among models with all possible combinations of climate covariates, and instead use Bayesian ridge regression to regulate, or constrain, the posterior distributions of each climate covariate [@Hooten2015].
Ridge regression is a specific application of statistical regularization that seeks to optimize model generality by trading off bias and variance.
As the name implies, statistical regularization involves the use of a regulator that constrains an optimization.
The natural regulator in a Bayesian application is the prior on the coefficient of interest.
Each of our statistical models includes the effects of climate covariates via the term $\textbf{z}'_t \boldsymbol{\beta}_c$ with prior $\boldsymbol{\beta}_c \sim \text{Normal}(\boldsymbol{\mu}_{\beta_c}, \sigma_{\beta_c}^2)$.
Since we standardized all climate covariates to have mean zero and variance one, we can set $\boldsymbol{\mu}_{\beta_c} = 0$ and let $\sigma_{\beta_c}^2$ serve as the regulator that can shrink covariates toward zero -- the smaller the prior variance, the more the posteriors of $\boldsymbol{\beta}_c$ are shrunk toward zero, and the stronger the penalty [@Hooten2015].

To find the optimal penalty (i.e., optimal value of the hyperparameter $\sigma_{\beta_c}^2$), we fit each statistical model with a range of values for $\sigma_{\beta_c}^2$ and compared predictive scores from leave-one-year-out cross-validation.
We performed the grid search over 24 values of $\sigma_{\beta_c}^2$, ranging from $\sigma_{\beta_c}^2 = 0.01$ to $\sigma_{\beta_c}^2 = 2.25$.
For each statistical model and each species we fit $13\times24=312$ models (13 years to leave out for cross-validation and 24 values of $\sigma_{\beta_c}^2$) -- a total of 4,992 models.
We calculated the log pointwise predictive density (*lppd*) to score each model's ability to predict the left-out data.
Thus, for training data $y_{train}$ and held-out data $y_{hold}$ at a given value of  $\sigma_\theta^2$ across all MCMC samples $s=1,2,...,S$ and all hold outs of data from year *t* to year *T*, and letting $\theta$ represent all unknowns, *lppd* is

@@

We chose the optimal prior variance for each species-statistical model combination as the one that produced the highest *lppd* and then fit each species-statistical model combination using the full data set for each species and the optimal prior variance.

### Population models
With the posterior distribution of the vital rate statistical models in hand, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to model populations based on individual-level data [@Ellner2006] and a quadrat-based version of an individually-based model (Quadrat-Based Model, QBM) to model populations based on quadrat-level data. We describe each in turn.

#### Integral projection model
We use a stochastic IPM [@Rees2009] that includes the climate covariates from the vital rate statistical models. In all simulations we ignore the random year effects so that interannual variation is driven solely by climate. We fit the random year effects in the vital rate regressions to avoid over-attributing variation to climate covariates. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size _v_ at time $t+1$ is

\vspace{-3em}
\begin{align}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{w_{j}}(u_{j}))n(u_{j},t)
\end{align}
\vspace{-3em}

@@

Since the IPM is spatially-implicit, we cannot calculate neighborhood crowding for specific genets ($w_{ij}$). Instead, we use an approximation ($\bar{w_{j}}$) that captures the essential features of neighborhood interactions [@Adler2010]. This approximation relies on a 'no-overlap' rule for conspecific genets to approximate the overdispersion of large genets in space [@Adler2010].

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=7in]{./components/climate_posteriors_figure.png}
  \caption{Posterior distributions of climate effects ($\boldsymbol{\beta}_C$) for each species and vital rate statistical model. Since our priors were constrained via ridge-regression, we highlight climate effects whose 80\% credible intervals do not overlap zero (red for negative coefficients, blue for positive coefficients). Kernel bandwidths of posterior densities were adjusted by a factor of 4 for visual clarity.}
\end{figure}

>>>>

\textsf{\textit{\footnotesize{\textsuperscript{1}Department of Wildland Resources and the Ecology Center, 5230 Old Main Hill, Utah State University, Logan, Utah 84322, USA; \textsuperscript{2}U.S. Geological Survey, Colorado Cooperative Fish and Wildlife Research Unit, Fort Collins, CO 80523, USA; \textsuperscript{3}Department of Fish, Wildlife, and Conservation Biology, Colorado State University, Fort Collins, CO 80523 USA}; \textsuperscript{4}Department of Statistics, Colorado State University, Fort Collins, CO 80523 USA}}

@@

Aggregate measures of population status, rather than individual performance, offer an intriguing alternative for modeling populations [@Clark2004; @Freckleton2011]. Population-level data cannot provide inference about demographic mechanisms, but might be sufficient for modeling future population states, especially because such data are feasible to collect across broad spatial extents [e.g., @Queenborough2011]. The choice between individual and population-level data involves a difficult trade-off: while individual-level data are necessary for mechanistic models, population-level data enable models that can be applied over greater spatial and temporal extents. An open question is how much forecasting skill is lost when we build models based on population rather than individual-level data.

@@

We compared the forecasting skill of statistical and population models based on aggregated, population-level data with the skill of models based on individual-level data. We used a demographic dataset that tracks the fates of individual plants from four species over 14 years to build two kinds of single-species population models, traditional models using individual growth, survival, and recruitment data and alternative models based on basal cover. We simulated the models to answer two questions motivated by the fact that the effects of intraspecific competition (density dependence) and interannual weather variability act at the level of the individual [@Clark2011b].
First, can population models fit using aggregated individual-level data (percent cover) adequately capture density dependence to produce forecasts as skillful as those from models fit to demographic data? 
Second, can population models fit using aggregated data adequately capture the influence of climate on population growth and, in turn, produce forecasts as skillful as those from models fit to demographic data?

Materials and Methods
---------------------
### Study site and data
Our demographic data come from a northern mixed grass prairie at the Fort Keogh Livestock and Range Research Laboratory near Miles City, Montana, USA ($46^{\circ}$ 19' N, $105^{\circ}$ 48' W). The dataset is available on Ecological Archives[^link] [@Anderson2011], and interested readers should refer to the metadata for a complete description. The site is 800 m above sea level and mean annual precipitation (1878-2009) is 334 mm, with most annual precipitation falling from April through September. The community is grass-dominated, and we focused on the four most abundant grass species: \emph{Bouteloua gracilis} (BOGR), \emph{Hesperostipa comata} (HECO), \emph{Pascopyrum smithii} (PASM), and \emph{Poa secunda} (POSE) (Fig. 1). *B. gracilis* is a warm-season perennial grass, whereas *H. comata*, *P. smithii*, and *Poa secunda* are cool-season perennial grasses. All species typically begin growth in the early spring, reach maximum growth and flower in early to mid summer (May-June), and disperse seed in mid to late summer (July-September). [LAST SENTENCES GLOSSES OVER LIKELY BIG DIFFERENCES IN PHENOLOGY. JUST DESCRIBE TIMING OF GROWING SEASON INSTEAD, TO SET UP CLIMATE COVARIATES?]

@@

From 1932 to 1945 individual plants were identified and mapped annually in 44 1-$\text{m}^2$ quadrats using a pantograph. The quadrats were distributed among six pastures, each assigned a grazing treatment of light (1.24 ha/animal unit month), moderate (0.92 ha/aum), and heavy (0.76 ha/aum) stocking rates (two pastures per treatment). In this analysis, we accounted for potential differences among the grazing treatments, but do not focus on grazing$\times$climate interactions. The annual maps of the quadrats were digitized and the fates of individual plants tracked and extracted using a computer program [@Lauenroth2008; @Chu2014a]. The permanent quadrats have not been relocated, but their distribution in six different pastures means the data represent a broad spatial distribution for the study area. Daily climate data are available for the duration of the data collection period (1932 - 1945) from the Miles City airport, Wiley Field, 9 km from the study site.

We modeled each grass population based on two levels of data: individual and quadrat. The individual data is the "raw" data. For the quadrat-level data we simply sum individual basal cover for each quadrat by species. This is equivalent to a near-perfect census of quadrat percent cover because measurement error at the individual-level is small [@Chu2014]. Based on these two datasets of 13 year-to-year transitions, we can compare population models built using individual-level data and aggregated, quadrat-level data. At the individual level, we explicitly model three vital rates: growth, survival, and recruitment. At the quadrat level, we model population growth as change in percent cover of quadrats with non-zero cover in year *t* and in year *t-1*, ignoring within-quadrat extirpation and colonization events because they are very rare in our time series ($N=16$ and $N=13$, respectively, across all species). Sample sizes for each species and vital rate model are shown in Table 1.

All R code and data necessary to reproduce our analysis is archived on GitHub as release v1.0[^revnote] (http://github.com/atredennick/MicroMesoForecast/releases). We have also deposited the v1.0 release on Dryad (*link here after acceptance*).

@@

### Stastical models of vital rates
At both levels of inference (individual and quadrat), the building blocks of our population models are vital rate regressions. For individual-level data, we fit regressions for survival, growth, and recruitment for each species. At the quadrat-level, we fit a single regression model for population growth. We describe the statistical models separately because they required different approaches. For both model types, we fit vital rate models with and without climate covariates. Models with climate effects contain five climate covariates that we chose \emph{a priori} based on previous model selection efforts using these data (Chu et al. *in press*) and expert advice (Lance Vermeire, *personal communication*): "water year" precipitation at \emph{t}-2 (lagppt); April through June precipitation at \emph{t}-1 and \emph{t} (ppt1 and ppt2, respectively) and April through June temperature at \emph{t}-1 and \emph{t} (TmeanSpr1 and TmeanSpr2, respectively), where \emph{t}-1 to *t* is the transition of interest. We also include interactions among same-year climate covariates (e.g., ppt1 $\times$ TmeansSpr1), resulting in a total of seven climate covariates.

We fit all models using a hierarchical Bayesian approach. We focus on the main process and the model likelihood in what follows (full model descriptions are in the Supporting Information). For the likelihood models, $\textbf{y}^X$ is always the relevant vector of observations for vital rate *X* (*X* = *S*, *G*, *R*, or *P*) for survival, growth, recruitment, or population growth). For example, $\textbf{y}^S$ is a vector of 0s and 1s indicating whether a genet survives from *t* to *t+1*, or not, for all observation years and quadrats. All model parameters are species-specific, but we omit subscripts for species in model descriptions below to reduce visual clutter. For brevity, we only describe models with climate covariates included, but models without climate covariates are simply the models described below with the climate effects removed.

#### Vital rate models at the individual level
We used logistic regression to model the probability that genet *i* in quadrat *q* survives from time *t* to *t*+1 ($s_{i,q,t}$):

\vspace{-3em}
\begin{align}
y_{i,q,t}^{S} &\sim \text{Bernoulli}(s_{i,q,t}) \\
\text{logit}(s_{i,q,t}) &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t})
\end{align}
\vspace{-3em}

where $x_{i,q,t}$ is the log of genet *i* basal area at time *t*, $\beta_{0,t}$ is a year specific intercept, $\beta_Q$ is the random effect of quadrat group location, $\beta_{s,t}$ is the year-specific slope parameter for size, **z** is a vector of *p* climate covariates specific to year *t*, $\boldsymbol{\beta}_c$ is a vector of fixed climate effects of length *p*, $\beta_{d,1}$ is the effect of intraspecific crowding experienced by the focal genet at time *t* ($w_{i,q,t}$), and $\beta_{d,2}$ is a size by crowding ($x_{i,q,t}w_{i,q,t}$) interaction effect.

@@

\begin{equation}
w_{i,q,t} = \sum_k e^{-\delta d_{ik,q,t}^{2}}u_{k,q,t}.
\end{equation}

In the above, $w_{i,t}$ is the crowding that genet _i_ in year _t_ experiences from conspecific neighbors in quadrat *q*. The spatial scale over which conspecific neighbors exert influence on any genet is determined by $\delta$. The function is applied for all _k_ conspecific genets that neighbor the focal genet at time _t_, and $d_{ik,q,t}$ is the distance between genet _i_ and conspecific genet _k_ in quadrat *q*. We use regression-specific (survival and growth) $\delta$ values estimated by @Chu2015.

We modeled growth as a Gaussian process describing log genet size ($y_{i,q,t+1}^{G}$) at time $t+1$ in quadrat *q* as a function of log size at time $t$ and climate covariates:

\vspace{-3em}
\begin{align}
y_{i,q,t+1}^G &\sim \text{Normal}(\mu_{i,q,t+1}, \sigma^2_{x_{i,q,t+1}}\textbf{I}) \\
\mu_{i,q,t+1} &= \beta_{0,t} + \beta_{s,t}x_{i,q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d,1} w_{i,q,t} + \beta_{d,2} (x_{i,q,t}w_{i,q,t})
\end{align}
\vspace{-3em}

where $\mu_{i,q,t+1}$ is log of genet *i*s predicted size at time *t* + 1, and all other parameters are as described for the survival regression. We capture non-constant error variance in growth by modeling the variance in the growth regression ($\sigma^2_{x{i,q,t+1}}$) as a nonlinear function of predicted genet size:

\vspace{-3em}
\begin{align}
\sigma^2_{x_{i,q,t+1}} = a \, \text{exp}[b \times \mu_{i,q,t+1}]
\end{align}
\vspace{-3em}

where $\mu_{i,q,t+1}$ is log of predicted genet size predicted from the growth regression (Eq. 4), and *a* and *b* are constants.

Our data allows us to track new recruits, but we cannot assign a specific parent to new genets. 
Therefore, we model recruitment at the quadrat level.
We assume the number of individuals, $y^{R}_{q,t+1}$, recruiting at time $t+1$ in quadrat *q* follows a negative binomial distribution:

@@

\vspace{-3em}
\begin{align}
\lambda_{q,t+1} = \tilde{c}_{q,t} \, \text{exp}\left(\beta_{0,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c + \beta_{d}\sqrt{\tilde{c}_{q,t}}\right)
\end{align}
\vspace{-3em}

where $\tilde{c}_{q,t}$ is effective cover ($\text{cm}^2$) of the focal species in quadrat *q* at time *t*, and all other terms are as in the survival and growth regressions.
Effective cover is a mixture of observed cover (*c*) in the focal quadrat (*q*) and the mean cover across the entire group ($\bar{c}$) of *Q* quadrats in which *q* is located:

\vspace{-3em}
\begin{align}
\tilde{c}_{q,t} = pc_{q,t} + (1-p)\bar{c}_{Q,t}
\end{align}
\vspace{-3em}

where $p$ is a mixing fraction between 0 and 1 that is estimated when fitting model.

#### Population model at the quadrat level
The statistical approach used to model aggregated data depends on the type of data collected. We have percent cover data, which can easily be transformed to proportion data in our case because plant areas were scaled by plot area. An obvious choice for fitting a linear model to proportion data is beta regression because the support of the beta distribution is (0,1), which does not include true zeros or ones. However, when we used fitted model parameters from a beta regression in a quadrat-based population model, the simulated population tended toward 100% cover for all species. We therefore chose a modeling approach based on a truncated log-normal likelihood. The model for quadrat cover change from time $t$ to $t+1$ is

\vspace{-3em}
\begin{align}
y_{q,t+1}^{P} &\sim \text{LogNormal}(\mu_{q,t+1}, \sigma^2)_{0}^{1} \\
\mu_{q,t+1} &= \beta_{0,t} + \beta_{s,t}x_{q,t} + \beta_Q + \textbf{z}'_t \boldsymbol{\beta}_c
\end{align}
\vspace{-3em}

where $\mu_{q,t+1}$ is the log of proportional cover in quadrat *q* at time $t+1$, and all other parameters are as in the individual-level growth model (Eq. 4) except that *x* now represents log of proportional cover. The log normal likelihood includes a truncation (subscript 0, superscript 1) to ensure that predicted values do not exceed 100% cover.

### Model fitting and statistical regularization
####  Model fitting
Our Bayesian approach to fitting the vital rate models required choosing appropriate priors for unknown parameters and deciding which, if any, of those priors should be hierarchical. We decided to fit models where all terms were fit by species. Within a species, we fit yearly size effects and yearly intercepts hierarchically, where year-specific coefficients were drawn from global distributions representing the mean size effect and intercept. Quadrat random effects were also fit hierarchically, with quadrat offsets being drawn from distributions with mean zero and a shared variance term (independent Gaussian priors, Appendix A). Climate effects were modeled as independent covariates whose prior distributions were optimized for prediction using statistical regularization (see **Statistical regularization: Bayesian ridge regression** below).

All of our analyses (model fitting and simulating) were conducted in R [@R2013]. We used the 'No-U-Turn' MCMC sampler in Stan [@stan2014]  to estimate the posterior distributions of model parameters using the package 'rstan' [@rstan2014]. We obtained posterior distributions for all model parameters from three parallel MCMC chains run for 1,000 iterations after discarding an initial 1,000 iterations. Such short MCMC chains may surprise readers more familiar with other MCMC samplers (i.e., JAGS or WinBUGS), but the Stan sampler reduces the number of iterations needed to achieve convergence. We assessed convergence visually and checked that scale reduction factors for all parameters were less than 1.1. For the purposes of including parameter uncertainty in our population models, we saved the final 1,000 iterations from each of the three MCMC chains to be used as randomly drawn values during population simulation. We report the posterior mean, standard deviation, and 95% Bayesian Credible Intervals for every parameter of each model for each species in the Supporting Information (Tables S#-S#).

#### Statistical regularization: Bayesian ridge regression
For models with climate covariates, our objective is to model the response of our focal grass species to interannual variation in climate, even if those responses are weak.
Therefore, we avoid selecting among models with all possible combinations of climate covariates, and instead use Bayesian ridge regression to regulate, or constrain, the posterior distributions of each climate covariate [@Gerber2015; @Hooten2015].
Ridge regression is a specific application of statistical regularization that seeks to optimize model generality by trading off bias and variance.
As the name implies, statistical regularization involves the use of a regulator that constrains an optimization.
The natural regulator in a Bayesian application is the prior on the coefficient of interest.
Each of our statistical models includes the effects of climate covariates via the term $\textbf{z}'_t \boldsymbol{\beta}_c$ with prior $\boldsymbol{\beta}_c \sim \text{Normal}(\boldsymbol{\mu}_{\beta_c}, \sigma_{\beta_c}^2)$.
Since we standardized all climate covariates to have mean zero and variance one, we can set $\boldsymbol{\mu}_{\beta_c} = 0$ and let $\sigma_{\beta_c}^2$ serve as the regulator that can shrink covariates toward zero -- the smaller the prior variance, the more the posteriors of $\boldsymbol{\beta}_c$ are shrunk toward zero, and the stronger the penalty [@Hooten2015].

To find the optimal penalty (i.e., optimal value of the hyperparameter $\sigma_{\beta_c}^2$), we fit each statistical model with a range of values for $\sigma_{\beta_c}^2$ and compared predictive scores from leave-one-year-out cross-validation.
We performed the grid search over 24 values of $\sigma_{\beta_c}^2$, ranging from $\sigma_{\beta_c}^2 = 0.01$ to $\sigma_{\beta_c}^2 = 2.25$.
For each statistical model and each species we fit $13\times24=312$ iterations of the model fitting algorithm to search $\sigma_{\beta_c}^2$ for the optimal value (13 years to leave out for cross-validation and 24 values of $\sigma_{\beta_c}^2$) -- a total of 4,992 model fits.
We calculated the log pointwise predictive density (*lppd*) to score each model's ability to predict the left-out data [@Gelman2014].
Thus, for training data $y_{train}$ and held-out data $y_{hold}$ at a given value of  $\sigma_\theta^2$ across all MCMC samples $s=1,2,...,S$ and all hold outs of data from year *t* to year *T*, and letting $\theta$ represent all unknowns, *lppd* is

@@

We chose the optimal prior variance for each species-statistical model combination as the one that produced the highest *lppd* and then fit each species-statistical model combination using the full data set for each species and the optimal prior variance.
We calculated the *lppd* from posterior samples using the algorithm from @Vehtari2016.

### Population models
Using samples from the posterior distribution of the vital rate statistical models, it is straightforward to simulate the population models. We used an Integral Projection Model (IPM) to simulate populations based on individual-level data [@Ellner2006] and a quadrat-based version of an individually-based model (Quadrat-Based Model, QBM) to simulate populations based on quadrat-level data. We describe each in turn.

#### Integral projection model
We use a stochastic IPM [@Rees2009] that includes the climate covariates from the vital rate statistical models. In all simulations, we ignore the random year effects so that interannual variation is driven solely by climate. We fit the random year effects in the vital rate regressions to avoid over-attributing variation to climate covariates. Our IPM follows the specification of @Chu2015 where the population of species _j_ is a density function $n(u_{j},t)$ giving the density of sized-_u_ genets at time _t_. Genet size is on the natural log scale, so that $n(u_{j},t)du$ is the number of genets whose area (on the arithmetic scale) is between $e^{u_{j}}$ and $e^{u_{j}+du}$. The density function for any size _v_ at time $t+1$ is

\vspace{-3em}
\begin{align}
n(v_{j},t+1) = \int_{L_{j}}^{U_{j}} k_{j}(v_{j},u_{j},\bar{w_{j}}(u_{j}))n(u_{j},t)du_{j}
\end{align}
\vspace{-3em}

@@

The IPM is spatially-implicit, thus we cannot calculate neighborhood crowding for specific genets ($w_{ij}$). Instead, we use an approximation ($\bar{w_{j}}$) that captures the essential features of neighborhood interactions [@Adler2010]. This approximation relies on a 'no-overlap' rule for conspecific genets to approximate the overdispersion of large genets in space [@Adler2010].

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=7in]{./components/climate_posteriors_figure.png}
  \caption{Posterior distributions of climate effects ($\boldsymbol{\beta}_C$) for each species and vital rate statistical model. Because our priors were constrained via ridge-regression, we highlight climate effects whose 80\% credible intervals do not overlap zero (red for negative coefficients, blue for positive coefficients). Kernel bandwidths of posterior densities were adjusted by a factor of 4 for visual clarity.}
\end{figure}
